---
title: "p8105_hw2_tt2714"
author: "Tiffany Tu"
date: "9/30/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(readxl)
library(stringr)
library(dplyr)
library(p8105.datasets)
library(knitr)
library(kableExtra)
options(scipen=999)
```

## Problem 1
#### Data import and cleaning 
This dataset contains information on New York City transit lines, station names, station latitude and longitude, routes served, type of entry, available entry, available vending machine, and ADA compliance. My data cleaning steps include cleaning all column names with janitor, selecting only wanted columns of the dataset, and converting the entry column from yes/no to logical variables.  
  
The dimension of the dataset is 1868 x 19. It is not tidy because there is too much overlapping information in the columns. It is unclear how many distinct transit lines, stations, or routes there are in the system, making it difficult to analyze and to find relevant information. The station latitude/longitude columns also give repetitive data. 
```{r data import 1, message=FALSE}
nyc_transit = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(-division, -exit_only, -staffing, -staff_hours, -ada_notes, -free_crossover:-entrance_location) %>% 
  replace(., is.na(.), "") %>% 
  unite(route, route1:route11, sep = " ") %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) %>% 
  mutate(vending = ifelse(vending == "YES", TRUE, FALSE)) 

head(nyc_transit)
```

#### Dataset overview: 
```{r}
station_nyc_transit = distinct(nyc_transit, station_name, route, .keep_all = TRUE)
Atrain_stations = station_nyc_transit[grep("A", station_nyc_transit$route), ]
```

There are `r nrow(station_nyc_transit)` distinct stations. `r length(which(station_nyc_transit$ada == TRUE))` of the stations are ADA compliant. The proportion of station entrances/exits without a vending that allows entrance is `r length(which(station_nyc_transit$vending == FALSE))/length(which(station_nyc_transit$entry == TRUE))`.  
There are `r nrow(Atrain_stations)` stations that serve the A train, of these stations `r length(which(Atrain_stations$ada == TRUE))` are ADA compliant.

## Problem 2
#### Mr. Trash Wheel dataset (2018-7)
```{r}
trashwheel = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = cell_cols(1:14)) %>% 
  janitor::clean_names() %>% 
  filter(str_detect(month, "Total") == FALSE) %>% 
  mutate_at(13, as.integer, 0)

head(trashwheel)
```

#### 2016 and 2017 precipitation data...
```{r clean & combine 2016/2017 dataset}
trashwheel_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  add_column(., year = 2017) %>% 
  select(1, 3, total.precipitation = 2) 

head(trashwheel_2017)

trashwheel_2016 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  add_column(., year = 2016) %>% 
  select(1, 3, total.precipitation = 2)

head(trashwheel_2016)

#combine 2016 and 2017 datasets
trashwheel_1617 = rbind(trashwheel_2016, trashwheel_2017) %>% 
  mutate(month.name[month], month = NULL) %>% 
  select(month = 3, 1, 2)

trashwheel_1617
```

#### Dataset overview

The Mr. Trash Wheel dataset provides a total of `r nrow(trashwheel)` dumpster observations from 2014 through 2018. For 2017, we are able to see `r length(which(trashwheel$year == 2017))` observations for the whole year. For 2016, we have `r length(which(trashwheel$year == 2016))` observations. There is an average of `r mean(subset(trashwheel, year == 2017)$cigarette_butts)` cigarette butts observed in 2017 as compared to an average of `r mean(subset(trashwheel, year == 2016)$cigarette_butts)` observed in 2016, which is a huge improvement. In 2018, the number of cigarette butts continued to decrease. So far we are seeing `r mean(subset(trashwheel, year == 2017)$cigarette_butts) - mean(subset(trashwheel, year == 2018)$cigarette_butts)` less cigarette butts from 2017. For the past five years, a median of `r median(trashwheel$weight_tons)` tons of weight and a median of `r median(trashwheel$volume_cubic_yards)` cubic yards in dumpsters were observed.

Total precipitation for 2017 is `r sum(trashwheel_2017[, 3])`. The median number of sports balls in a dumpster in 2016 is `r median(subset(trashwheel, year == 2016)$sports_balls, na.rm = TRUE)`.

Here we have a quick table that provides the number of observations per year for reference.
```{r echo=FALSE}
trashwheel %>% 
  group_by(year) %>%
  count() %>%
  t() %>% 
  kable()
```

## Problem 3
Dataset from Behavioral Risk Factors Surveillance System for Selected Metropolitan Area Risk Trends is uploaded and cleaned for analysis on "Overall Health" topic.

```{r clean BRFSS data}
overallhealth_brfss_smart2010 = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  subset(., topic == "Overall Health") %>% 
  select(., -class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location)) %>%
  spread(response, data_value) %>%
  janitor::clean_names() %>% 
  mutate(., excellent_or_verygood = rowSums(cbind(excellent, very_good)))

head(overallhealth_brfss_smart2010)
```

There are `r length(unique(overallhealth_brfss_smart2010$locationabbr))` unique locations in this dataset. Every state is included in this dataset, including District of Columbia. The most frequent occurence is `r max(overallhealth_brfss_smart2010$locationabbr)` out of all the states represented.  

In 2002, the median numeber of "Excellent" response value is `r median(subset(overallhealth_brfss_smart2010, year == 2002)$excellent, na.rm = TRUE)`.

#### "Excellent" responses in 2002
```{r histogram}
ggplot(subset(overallhealth_brfss_smart2010, year == 2002), aes(x = excellent)) + 
  geom_histogram(binwidth = 2) + labs(title = "'Excellent' response values in 2002", x = "proportion of 'excellent' response")
```

#### "Excellent" responses in New York County and Queens County
```{r scatterplot}
nycounty = overallhealth_brfss_smart2010 %>% 
  filter(locationdesc == "NY - New York County" | locationdesc == "NY - Queens County")

ggplot(nycounty, aes(x = year, y = excellent)) + geom_point(aes(color = locationdesc)) + 
  labs(x = "year", y = "proportion of 'excellent' response")
```
